% !TEX TS-program = xelatex
% Compile with: xelatex or lualatex

\documentclass[11pt,a4paper,sans]{moderncv}

% ModernCV style
\moderncvstyle{classic}    % casual, classic, banking, oldstyle, fancy
\moderncvcolor{blue}       % blue, orange, green, red, purple, grey, black

% Encoding and layout
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[scale=0.85]{geometry}
\setlength{\hintscolumnwidth}{3cm}

%----------------------------------------------------------------------------------
% Personal data
%----------------------------------------------------------------------------------
\name{Julien}{Soulé}
\title{Doctor in Computer Science}
\address{35 Rue Mathieu-de-la-Drôme}{26000 Valence, France}{}
\phone[mobile]{+33~6~77~63~12~13}
\email{julien.soule@hotmail.fr}
\homepage{julien6.github.io/home/}
% \social[linkedin]{julien-soulé-6b2b27173}
% \social[github]{julien6}

\begin{document}

%----------------------------------------------------------------------------------
% Letter
%----------------------------------------------------------------------------------
\recipient{Helsing}{Recruitment Team}
\date{\today}
\opening{Dear Hiring Team,}
\closing{Sincerely,}
\enclosure[Attachment]{Curriculum Vitae}

\makelettertitle

I am writing to apply for the position of \textit{AI Research Engineer in Reinforcement Learning}. I recently completed a PhD in Computer Science at Université Grenoble Alpes, conducted in collaboration with Thales Land \& Air Systems. My research lies primararly at the intersection of \textbf{Multi-Agent Reinforcement Learning (MARL)} and \textbf{organizational modeling} with application for \textbf{cyberdefense} centered on the \href{https://link.springer.com/book/10.1007/978-3-031-29269-9}{\textbf{\textbf{AICA}}} (Autonomous Intelligent Cyberdefense Agent) initiative. My thesis makes part of the work carried out by the \textbf{NATO IST-152} working group and its successor, the \href{https://www.aica-iwg.org/}{\textbf{AICA IWG}} (AICA International Working Group), where I serve as treasurer. This involvement has exposed me to a rich, international community of researchers and practitioners working on AI-driven attack surface assessment and autonomous defensive strategies, providing both inspiration and practical guidance for my research agenda in applying \textbf{RL and MARL techniques} to cyberdefense and beyond. Throughout \href{https://julien6.github.io/home/#thesis}{\textbf{my PhD}} at \textit{Université Grenoble Alpes} (UGA) in collaboration with \textit{Thales LAS}, I have approached the design of a \textbf{Multi-Agent System} (MAS) cyberdefense as a sequential decision-making problem under uncertainty, where autonomous agents must explore, monitor, and protect complex, heterogeneous infrastructures.

My academic background is rooted in computer engineering and cybersecurity, and my research progressively converged toward the question of how collections of learning agents can be structured, constrained, and evaluated as coherent systems rather than as isolated learners. During my doctoral work, I explored the use of organizational models to guide multi-agent reinforcement learning, with the objective of improving coordination, robustness, and interpretability. This research led to several peer-reviewed publications and to the development of research software platforms combining simulation, learning, and system-level analysis.

In parallel, my experience in industry, notably at Thales and Atos, allowed me to work on complex and safety-critical systems in operational contexts. These experiences strongly shaped my approach to research, with particular attention to system architecture, experimental reproducibility, and the practical constraints that arise when learning-based components are embedded in real-world systems. Over time, I have come to view research and engineering not as separate activities, but as two complementary aspects of building reliable autonomous systems.

At this stage of my career, I am looking to continue working on problems related to autonomous decision-making, learning under uncertainty, and the coordination of complex systems, within an environment where long-term technical depth and real-world impact matter. I am particularly interested in settings that combine advanced AI research with demanding operational constraints.

I would be glad to discuss my background and interests in more detail, and to explore whether my experience could be a good fit for your team.

\makeletterclosing

\end{document}
