%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Remerciements}{Remerciements}

% \begin{flushright}{\slshape    
%     We have seen that computer programming is an art, \\ 
%     because it applies accumulated knowledge to the world, \\ 
%     because it requires skill and ingenuity, and especially \\
%     because it produces objects of beauty.} \\ \medskip
%     --- \defcitealias{knuth:1974}{Donald E. Knuth}\citetalias{knuth:1974} \citep{knuth:1974}
% \end{flushright}



% \bigskip

% \begingroup
% \let\clearpage\relax
% \let\cleardoublepage\relax
% \let\cleardoublepage\relax
% \chapter*{Remerciements}

% Faire un doctorat peut sembler être un chemin solitaire, mais je ne l'ai certainement pas parcouru seul.

% Je tiens tout d'abord à remercier mes directeurs de thèse, Jean-Paul et Michel ainsi qu'à Louis-Marie qui m'a accompagné au sein de Thales et Paul qui m'a suivi tout au long de ces trois années. Ils m'ont guidé tout au long de ce processus, m'ont permis de découvrir ce qu'est la recherche (y compris les erreurs) mais aussi d'intervenir lorsque c'était nécessaire. Ils ont toujours été là pour moi, même pendant leurs vacances, en prenant le temps de relire mon manuscrit de thèse.

% Un grand merci également aux membres du jury de thèse. Je tiens à remercier les rapporteurs de ma thèse: Laurent Vercouter et Gauthier Picard pour leurs relectures de mon manuscrit. Je remercie également les examinateurs: Aurélie Beynier pour avoir assumé le rôle de présidente du jury, Flavien Balbo pour m'avoir également suivi dans le cadre du CSI, Laeticia Matignon et Oum-el-Kheir Aktouf de bien avoir accepté d'assister et examiner ma soutenance de thèse.

% J'ai rencontré de nombreux étudiants incroyables au cours de mon séjour au laboratoire. Je voudrais surtout mentionner Arthur, Vincent, Thinh et Karem pour m'avoir accueilli. Je tiens à remercier Minh Tuan, Sébastien et Maximilian pour leur gentillesse et pour tous les échanges que nous avons eus.

% Je remercie également François Suro et Simon Gay, qui m'ont donné régulièrement des retours et m'ont permis de garder le cap quand j'étais motivée mais manquais encore de confiance en mon travail.

% Un grand merci à Clément Raïevsky, Laurent Lefèvre, Annabelle Mercier et Louis Morge-Rollet pour m'avoir aidé dans mes expériences d'enseignement. Travailler au LCIS n'aurait pas été aussi facile sans Patricia et Carole qui m'ont aidé pour affronter la montagne de formalités administratives, ainsi qu'Arthur, Marie et Nicolas qui m'ont permis d'utiliser les moyens techniques dont j'avais besoin.

% Je tiens également à remercier mes amis Dorian et Thomas, qui sont également sur ce chemin du doctorat, ainsi que Joaquin, qui a toujours été là pour m'aider à faire une pause dans mes réflexions sur mes études.

% Et enfin, je dois tout à ma famille, qui m'a soutenue sans réserve tout au long de cette aventure. Rien de tout cela n'aurait été possible sans eux.

% \endgroup



\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}


Pursuing a doctorate may seem like a solitary path, but I certainly didn't walk it alone.

First and foremost, I would like to thank my thesis supervisors, Jean-Paul and Michel, as well as Louis-Marie, who supported me at Thales, and Paul, who guided me throughout these three years. They guided me throughout this process, allowing me to discover what research is all about (including the mistakes), and also to intervene when necessary. They were always there for me, even during their holidays, taking the time to proofread my thesis manuscript.

A big thank you also to the members of the thesis committee. I would like to thank my thesis reviewers, Laurent Vercouter and Gauthier Picard, for their proofreading of my manuscript. I would also like to thank the examiners: Aurélie Beynier for serving as jury president, Flavien Balbo for his support within the CSI program, and Laeticia Matignon and Oum-el-Kheir Aktouf for agreeing to attend and examine my thesis defense.

I met many incredible students during my time at the lab. I would especially like to mention Arthur, Vincent, Thinh, and Karem for welcoming me. I would also like to thank Minh Tuan, Sébastien, and Maximilian for their kindness and for all the conversations we had.

I would also like to thank François Suro and Simon Gay, who regularly provided feedback and helped me stay on track when I was motivated but still lacked confidence in my work.

A big thank you to Clément Raïevsky, Laurent Lefèvre, Annabelle Mercier, and Louis Morge-Rollet for their support during my teaching experiences. Working at LCIS wouldn't have been so easy without Patricia and Carole, who helped me navigate the mountain of paperwork, and Arthur, Marie, and Nicolas, who provided me with the technical resources I needed.

I also want to thank my friends Dorian and Thomas, who are also pursuing their doctorates, and Joaquin, who has always been there to help me take a break from my studies.

And finally, I owe everything to my family, who have supported me wholeheartedly throughout this adventure. None of this would have been possible without them.

\endgroup
