\documentclass[10pt]{beamer}
\usetheme{Madrid}
\usecolortheme{default}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{ragged2e}
\hypersetup{colorlinks=true,linkcolor=blue,urlcolor=blue}

\title[Model-Based RL with Structure and Safety]{From Multi-Agent Reinforcement Learning to\\Structured and Verifiable World Models}
\subtitle{Discussion for the Post-Doctoral Position -- Télécom Paris (IMT)}
\author{Julien Soulé}
\institute{Université Grenoble Alpes / Thales LAS}
\date{Interview -- November 2025}

% --------------------------------------------------------
\begin{document}

\begin{frame}
    \titlepage
\end{frame}

% --------------------------------------------------------
\begin{frame}{Outline}
    \tableofcontents
\end{frame}

% --------------------------------------------------------
\section{Background and Motivation}
\begin{frame}{Background and Motivation}
    \justifying
    \begin{itemize}
        \item Research background: Multi-Agent Reinforcement Learning (MARL), World Models, hybrid symbolic-learning architectures.
        \item Motivation: Current learned models are powerful but brittle, opaque, and hard to verify.
        \item Postdoc goal: Develop \textbf{structured and verifiable world models} for safety-critical domains.
    \end{itemize}
\end{frame}

% --------------------------------------------------------
\section{Past Research Contributions}
\begin{frame}{PhD Highlights}
    \justifying
    \textbf{MAMAD Framework}:Modeling, Training, Analyzing, and Transferring.
    \vspace{0.2cm}
    \begin{itemize}
        \item World models used as digital twins for simulation.
        \item Combines MARL, simulation via learned environment dynamics, and unsupervised behaviour analysis.
        \item Integrates organizational roles/goals ($\mathcal{M}OISE^+$) for explainable and controllable learning.
        \item Implemented in the \href{https://github.com/julien6/CybMASDE}{\textbf{CybMASDE}} platform (automated modeling-training-analysis-transferring pipeline).
    \end{itemize}
    \vspace{0.2cm}

    \vspace{0.2cm}
    \textbf{Case studies}
    \begin{itemize}
        \item \textbf{Academic research games}: Overcooked-AI, Warehouse management, Predator-prey\dots.
        \item \textbf{Industrial case study -- Kubernetes cluster pod orchestration} with Thales LAS: RL-based autoscaling and load balancing.
    \end{itemize}

    \vspace{0.2cm}
    \footnotesize
    \textit{References:} Soulé et al., AIAI 2024; AAMAS 2025; IEEE CLOUD 2025; JAAMAS (under revision).
\end{frame}

% --------------------------------------------------------
\section{Research Fit with Télécom Paris}
\begin{frame}{Alignment with the Postdoc Topics}
    \justifying
    \textbf{Structured / Neurosymbolic World Models}
    \begin{itemize}
        \item Integrate sub-symbolic dynamics with explicit symbolic or physical constraints.
        \item Link to Mohan, Zhang \& Lindauer (2024) -- \textit{Structure in Deep RL}.
    \end{itemize}
    \vspace{0.2cm}
    \textbf{Compositional Reasoning for Safety}
    \begin{itemize}
        \item Modular world models (perception, dynamics, constraint modules).
        \item Inspired by Bakirtzis et al. (2025) -- \textit{Categorical Semantics of Compositional RL}.
    \end{itemize}
\end{frame}

% --------------------------------------------------------
\section{Proposed Research Direction}
\begin{frame}{Vision: Composable and Verifiable World Models}
    \justifying
    \textbf{Core idea:} Split the world model into modules with typed interfaces.
    \begin{block}{Modules}
        \begin{itemize}
            \item \textbf{Perception:} latent observation encoder + OOD detection.
            \item \textbf{Dynamics:} latent transition model (RSSM-like) + uncertainty estimation.
            \item \textbf{Constraints:} differentiable logic or hard safety guards.
        \end{itemize}
    \end{block}
    \textbf{Planning:} MPC / policy trained on model rollouts with constraint enforcement.
\end{frame}

% --------------------------------------------------------
\begin{frame}{Two Short-Term Research Axes}
    \justifying
    \begin{enumerate}
        \item \textbf{Constraint-Aware Dreamer:}
              \begin{itemize}
                  \item Integrate safety constraints and logic layers into latent model.
                  \item Evaluate on Safety-Gymnasium / DMControl.
              \end{itemize}
        \item \textbf{Graph-Constraint World Models:}
              \begin{itemize}
                  \item Represent dynamics as Graph Neural Network.
                  \item Constraints as graph predicates for compositional verification.
              \end{itemize}
    \end{enumerate}
    \vspace{0.2cm}
    Deliverables: open-source code, safety metrics, publication at NeurIPS/ICLR/ICML.
\end{frame}

% --------------------------------------------------------
\section{Discussion Points}
\begin{frame}{Questions and Collaboration Opportunities}
    \justifying
    \begin{itemize}
        \item Which formalisms for safety? (CMDPs, barrier functions, temporal logic\dots)
        \item When to use and combine training from explicit model (MPC) and policy learning from imagination?
        \item Possible industrial scenarios via the “Architecture of Complex Systems” chair.
        \item Opportunities to merge category-theoretic composition with engineering implementations.
    \end{itemize}
\end{frame}

% --------------------------------------------------------
\section{Conclusion}
\begin{frame}{Take-Home Message}
    \justifying
    \begin{itemize}
        \item My work bridges \textbf{multi-agent RL, world models, and organizational structure}.
        \item This postdoc is a natural next step toward \textbf{structured, safe, and interpretable MBRL}.
        \item I aim to include multi-agent scenarios (Dec-POMDP)
              \begin{itemize}
                  \item How to "reconstruct" the most likely world state from partial observations, make it explainable, visually renderable\dots
              \end{itemize}
        \item I bring practical implementation experience (RLlib, JAX/JaxMARL, Torch, Gym/PettingZoo, etc.) and a research plan.
    \end{itemize}
    \vspace{0.3cm}
    \centering
    \textbf{Thank you for your time!}
\end{frame}

% --------------------------------------------------------
\end{document}
