\section*{NeurIPS Paper Checklist}

\begin{enumerate}

      \item {\bf Claims}
      \item[] Question: Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
      \item[] Answer: \answerYes{}
      \item[] Justification: The abstract and introduction state the paper's contributions, assumptions, and empirical scope consistently with the proposed NS-MAWM framework, theoretical analyses, and experimental evaluation protocols presented in the main paper and appendices.

      \item {\bf Limitations}
      \item[] Question: Does the paper discuss the limitations of the work performed by the authors?
      \item[] Answer: \answerYes{}
      \item[] Justification: The paper discusses limitations related to symbolic-rule correctness, symbolic coverage assumptions, scalability, residual-dynamics modeling, computational cost, robustness under distribution shifts, and the non-guaranteed nature of the theoretical proof sketches.

      \item {\bf Theory assumptions and proofs}
      \item[] Question: For each theoretical result, does the paper provide the full set of assumptions and a complete (and correct) proof?
      \item[] Answer: \answerYes{}
      \item[] Justification: The paper provides lightweight theoretical analyses and proof sketches under simplifying assumptions to support the main intuitions behind the proposed framework. These analyses are not intended as complete guarantees for arbitrary practical settings, and their assumptions and limitations are discussed in the appendix.

      \item {\bf Experimental result reproducibility}
      \item[] Question: Does the paper fully disclose all the information needed to reproduce the main experimental results of the paper to the extent that it affects the main claims and/or conclusions of the paper (regardless of whether the code and data are provided or not)?
      \item[] Answer: \answerYes{}
      \item[] Justification: The paper details the baseline taxonomy, training and evaluation protocols, environments, symbolic-rule construction process, metrics, ablation methodology, and evaluation procedures in both the main paper and appendices.

      \item {\bf Open access to data and code}
      \item[] Question: Does the paper provide open access to the data and code, with sufficient instructions to faithfully reproduce the main experimental results, as described in supplemental material?
      \item[] Answer: \answerYes{}
      \item[] Justification: The paper describes the implementation framework, symbolic-rule inventory, and reproduction procedures. The authors intend to release anonymized code, configuration files, and experimental scripts upon publication.

      \item {\bf Experimental setting/details}
      \item[] Question: Does the paper specify all the training and test details (e.g., data splits, hyperparameters, how they were chosen, type of optimizer) necessary to understand the results?
      \item[] Answer: \answerYes{}
      \item[] Justification: Training procedures, evaluation regimes, baseline dimensions, optimization protocols, symbolic coverage configurations, rollout horizons, derived metrics, and evaluation procedures are described in the main paper and appendices.

      \item {\bf Experiment statistical significance}
      \item[] Question: Does the paper report error bars suitably and correctly defined or other appropriate information about the statistical significance of the experiments?
      \item[] Answer: \answerYes{}
      \item[] Justification: Results are aggregated across multiple seeds, variants, and repeated runs. The paper reports mean $\pm$ standard error and discusses variance-based robustness metrics throughout the evaluation protocol and full-results appendix.

      \item {\bf Experiments compute resources}
      \item[] Question: For each experiment, does the paper provide sufficient information on the computer resources (type of compute workers, memory, time of execution) needed to reproduce the experiments?
      \item[] Answer: \answerYes{}
      \item[] Justification: The paper reports resource-oriented metrics such as CPU usage, RAM usage, WM update time, and step time. Additional hardware-specific execution details and implementation artifacts will be included in the released code repository.

      \item {\bf Code of ethics}
      \item[] Question: Does the research conducted in the paper conform, in every respect, with the NeurIPS Code of Ethics \url{https://neurips.cc/public/EthicsGuidelines}?
      \item[] Answer: \answerYes{}
      \item[] Justification: The research follows standard ML research practices and complies with the NeurIPS Code of Ethics. The work focuses on general multi-agent world-model methodology rather than deployment-specific autonomous systems.

      \item {\bf Broader impacts}
      \item[] Question: Does the paper discuss both potential positive societal impacts and negative societal impacts of the work performed?
      \item[] Answer: \answerYes{}
      \item[] Justification: The paper discusses positive impacts related to interpretable and sample-efficient world models, as well as risks associated with autonomous decision-making systems, robustness failures, symbolic-rule biases, and potential misuse of model-based planning systems.

      \item {\bf Safeguards}
      \item[] Question: Does the paper describe safeguards that have been put in place for responsible release of data or models that have a high risk for misuse (e.g., pre-trained language models, image generators, or scraped datasets)?
      \item[] Answer: \answerNA{}
      \item[] Justification: The paper does not release high-risk generative models, scraped datasets, or deployment-ready autonomous systems requiring specific safeguards.

      \item {\bf Licenses for existing assets}
      \item[] Question: Are the creators or original owners of assets (e.g., code, data, models), used in the paper, properly credited and are the license and terms of use explicitly mentioned and properly respected?
      \item[] Answer: \answerYes{}
      \item[] Justification: Existing environments, frameworks, datasets, and baseline methods are cited and credited throughout the paper and appendices, including GridCraft, Overcooked, SMACv2, MAPPO, QMIX, Dreamer-style architectures, and related neuro-symbolic approaches.

      \item {\bf New assets}
      \item[] Question: Are new assets introduced in the paper well documented and is the documentation provided alongside the assets?
      \item[] Answer: \answerYes{}
      \item[] Justification: The paper documents the proposed NS-MAWM framework, symbolic-rule inventories, evaluation protocols, and experimental methodology. Additional implementation documentation and reproducibility artifacts will accompany the public release.

      \item {\bf Crowdsourcing and research with human subjects}
      \item[] Question: For crowdsourcing experiments and research with human subjects, does the paper include the full text of instructions given to participants and screenshots, if applicable, as well as details about compensation (if any)?
      \item[] Answer: \answerNA{}
      \item[] Justification: The paper does not involve crowdsourcing or research with human subjects.

      \item {\bf Institutional review board (IRB) approvals or equivalent for research with human subjects}
      \item[] Question: Does the paper describe potential risks incurred by study participants, whether such risks were disclosed to the subjects, and whether Institutional Review Board (IRB) approvals (or an equivalent approval/review based on the requirements of your country or institution) were obtained?
      \item[] Answer: \answerNA{}
      \item[] Justification: The paper does not involve human subjects or crowdsourcing experiments.

      \item {\bf Declaration of LLM usage}
      \item[] Question: Does the paper describe the usage of LLMs if it is an important, original, or non-standard component of the core methods in this research?
      \item[] Answer: \answerYes{}
      \item[] Justification: The appendix describes the use of LLM-assisted symbolic-rule generation to initialize partial symbolic-rule inventories before manual refinement and validation. The paper explicitly distinguishes automatically generated rules from manually refined symbolic rules.

\end{enumerate}