\documentclass[11pt]{letter}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{url}

\signature{Julien Soulé}
\address{Julien Soulé \\ Thales LAS / LCIS Laboratory \\ Valence, France \\ \texttt{julien.soule@univ-grenoble-alpes.fr}}

\begin{document}

\begin{letter}{Editor-in-Chief \\ Journal of Autonomous Agents and Multi-Agent Systems (JAAMAS)}

\opening{Dear Editor,}

Please find enclosed our manuscript entitled \textbf{“Assisting Multi-Agent System Design with MOISE+ and MARL: The MAMAD Method”}, which we submit for consideration as an original research article in the \emph{Journal of Autonomous Agents and Multi-Agent Systems (JAAMAS)}.

This work builds upon our AAMAS 2025 conference paper, entitled \textit{“An Organizationally-Oriented Approach to Enhancing Explainability and Control in Multi-Agent Reinforcement Learning”}, in which we introduced MOISE+MARL -- a framework that integrates organizational modeling with Multi-Agent Reinforcement Learning (MARL).
In this extended manuscript, we embed this contribution into a broader method called \textbf{MAMAD} (MOISE+MARL Assisted MAS Design), which aims to automate the design of Multi-Agent Systems (MAS) through a structured four-phase pipeline: Modeling, Training, Analyzing, and Transferring.

The paper addresses four key challenges still prevalent in the literature: the integration of MARL within AOSE practices; understanding emergent agent behaviors; guiding policy learning through organizational constraints; and automating the full MAS design process. Our results show that MAMAD leads to more structured, compliant, and explainable agent behaviors across several environments and MARL algorithms, with limited human intervention.

We believe this work will be of interest to JAAMAS readers because it bridges the gap between symbolic organizational models and data-driven learning approaches. It offers both theoretical formalization and practical implementation via our open-source tool \texttt{CybMASDE}, making it relevant for both researchers and practitioners in multi-agent systems, reinforcement learning, and organizational AI.

We confirm that this manuscript has not been published elsewhere and is not under consideration by another journal. All authors have approved the manuscript and agree with its submission to JAAMAS. The corresponding author is Julien Soulé (\texttt{julien.soule@univ-grenoble-alpes.fr}).

Thank you for considering our submission. We remain at your disposal for any further information.

\closing{Sincerely,}

\end{letter}

\end{document}
